---
title: "Classe Prediction Project"
author: "K Johnson"
date: '`r Sys.Date()`'
output: html_document
---

###Executive Summary

This project's goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants and predict and quantify how well they do it. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: 

1. exactly according to the specification (Class A), 
2. throwing the elbows to the front (Class B), 
3. lifting the dumbbell only halfway (Class C), 
4. lowering the dumbbell only halfway (Class D) and 
5. throwing the hips to the front (Class E).

Read more: http://groupware.les.inf.puc-rio.br/har#sbia_paper_section#ixzz46tAmwNTy

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har

The following files have been provided for this project:

1. pml-training.CSV
2. pml-testing.CSV

###Exploratory Analysis



```{r ReadFiles, echo=FALSE,include=FALSE,message=FALSE}
##install.packages("caret")
library(caret)
setwd("C:/Users/FIS/Documents/Coursera")
trainingP <- read.csv(file="pml-training.CSV",head=TRUE,sep=",")
testingP <- read.csv(file="pml-testing.CSV",head=TRUE,sep=",")
str(trainingP)
inTrain <- createDataPartition(y=trainingP$classe,p=0.75, list=FALSE)
training <- trainingP[inTrain,]
testing <- trainingP[-inTrain,]
```
A display of the class and some of the values in each column in the training dataset is shown in Appendix, Exhibit 1. The number of observations is `r nrow(trainingP)`.

###Regression Models

```{r RegMod1, echo=FALSE,include=FALSE,cache=TRUE,message=FALSE}
model1 <- train(classe ~ total_accel_arm + total_accel_belt + total_accel_dumbbell + total_accel_forearm + num_window,preProcess=c("center", "scale"),data=training)
print(model1)
###Selection of Model
```
```{r RegMod2, echo=FALSE,include=FALSE,cache=TRUE,message=FALSE}
model2 <- train(classe ~ total_accel_arm + total_accel_belt + total_accel_dumbbell + total_accel_forearm ,preProcess=c("center", "scale"),data=training)
print(model2)
###Selection of Model
```
Looking at the data, the model use these  predictors:
    1. total_accel_arm 
    2. total_accel_belt
    3. total_accel_dumbbell 
    4. total_accel_forearm
    5. num_window
    
    
The model selection has accuracy of .998. The print out of the models are shown in Exhibit 2 in the Appendix.

###K fold Cross Validation
```{r ReSamp, echo=FALSE,include=FALSE,cache=TRUE,warning=FALSE}
train_control <- trainControl(method="cv", number=10)
modelx <- train(classe ~ total_accel_arm + total_accel_belt + total_accel_dumbbell + total_accel_forearm + num_window,trControl=train_control,method="nb",data=training)
print(modelx)
```
This analysis used 10 fold cross validation to construct Naive Bayes model using the  selected predictors (from model 1). The results are shown in Exhibit 3.

## Predict Classe using the provided testing file and the selected model, The results are shown in Exhibit 4.
```{r TestPred, echo=FALSE,include=FALSE,cache=TRUE,warning=FALSE}
predict1 <- predict(model1,newdata=testingP)
print(predict1)
```
Exhibit 5 shows the expected out of sample error using confusion matrix with testing data subset of training
### determine out of sample error using the confusion matrix.
```{r SampErr, echo=FALSE,include=FALSE,cache=TRUE,warning=FALSE}
predict2 <- predict(model1,newdata=testing)
confusionMatrix(testing$classe, predict2)
```
### APPENDIX

#### Figure 1 - Class and sample data from each field in the training dataset
```{r ref.label="ReadFiles",include=TRUE,echo=FALSE,message=FALSE}
```
#### Figure 2 -  models summary
#### model 1 = classe ~ total_accel_arm + total_accel_belt + total_accel_dumbbell + total_accel_forearm + num_window
```{r ref.label="RegMod1",include=TRUE,echo=FALSE,message=FALSE}
```
#### model 2 = classe ~ total_accel_arm + total_accel_belt + total_accel_dumbbell + total_accel_forearm
```{r ref.label="RegMod2",include=TRUE,echo=FALSE,message=FALSE}
```
#### Figure 3 - K fold cross validation results on selected model
```{r ref.label="ReSamp",include=TRUE,echo=FALSE,warning=FALSE}
```
####  Figure 4 - Predictions using the Testing file
```{r ref.label="TestPred",include=TRUE,echo=FALSE,warning=FALSE}
```
#### Figure 5 - Out of sample error
```{r ref.label="SampErr",include=TRUE,echo=FALSE,warning=FALSE}
```
